---
title: "Data Viz Mini-Project"
author: "Matthew Huber"
date: "September 26, 2018"
output:
  word_document: default
  html_notebook: default
---

```{r setup, results='hide',warning=FALSE}
#Load tidyverse in order to use readr

library(tidyverse)
library(maps)
```

In order to explore the data for this assignment, I first had to read it in as a CSV file. I used the readr function read_csv for the data set itself, but I also read in the data using read.csv to compare to the read_csv file to see if there were any differences between the two that would affect the data being read in.

```{r}
#Read CSV into R

shooting <- readr::read_csv("/Users/matt/Documents/Class Docs/Data Science/Mini Project 1/fatal-police-shootings-data.csv")

#Brief look at the data set

head(shooting)
str(shooting)

#Read in as read.csv to compare

shootcsv <- read.csv("/Users/matt/Documents/Class Docs/Data Science/Mini Project 1/fatal-police-shootings-data.csv")

head(shootcsv)
str(shootcsv)

```

After reading in the data, I began my exploration using the following steps:

1) Exploration of Individual Variables

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_histogram(mapping=aes(x=date),color="white",binwidth=40) + ggtitle("Plot of Date of Death")
```

Date of Death - To me, the data seems fairly evenly distributed across the years that the data base is pulling for, with numberds jumping slightly in the summer months, which to me seems logical because when the weather gets nicer more people tend to be out and about committing crimes. Personally, if I were to cmmit a crime, I would rather do it when it's nice out than when it's cold or snowing.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=manner_of_death)) + geom_text(stat='count', aes(x=manner_of_death,label=..count..), vjust=-1) + ggtitle("Plot of Manner of Death")
```

Manner of Death - To me this variable doesn't quite make sense for inclusion. In the end, everyone was shot and only about 200 people were also tasered out of the almost 4000 victims since 2015.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=armed),color="white") + geom_text(stat='count', aes(x=armed,label=..count..), vjust=-1) + ggtitle("Plot of Weapons Used")
```

Armed - This variable is a mess. There are way too many categories for this plot to make any sense, and the fact that the data is so spread out, with many categories only having one observation, do little to add to the understanding of the data. I wanted to try and reclassify a lot of these values into an "other"" category and minimize the number of classes of weaponry. For example, guns, knives, handheld objects (baseball bats, hammers, etc), unarmed, other, etc. I decided to use the threshold of 100 as a minimum number of observations for a category to remain as is. Every class with less than 100 observations will be lumped together as other.

```{r fig.height=10, fig.width=10}
w = as.data.frame(table(shooting$armed))
names(w)[1]='armed'
shooting2 <- merge(shooting, w)

shooting2$armed <- ifelse(shooting2$Freq<=100,"other",shooting2$armed)

ggplot (data=shooting2) + geom_bar(mapping=aes(x=armed),color="white") + geom_text(stat='count', aes(x=armed,label=..count..), vjust=-1) + ggtitle("Plot of Weapons Used")
```

The amount of victims that were carrying a gun are still staggering, but now we can see more clearly that the other classes worth noting are knife, unarmed/undetermined, and toy weapon. The relationship between this and other variables would be very interesting to explore.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_histogram(mapping=aes(x=age),color="white",binwidth=10) + ggtitle("Plot of Age")
```

Age - The data appears to be a bit skewed, having a larger concentration of younger victims, which makes sense seeing as older individuals to me seem less likely to be committing crimes. I did find it interesting that there were about 144 observations without an age listed.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=gender),color="white") + geom_text(stat='count', aes(x=gender,label=..count..), vjust=-1) + ggtitle("Plot of Gender")
```

Gender - I am surprised that there were only 164 females and 3 individuals with no gender listed. It would be interesting to see the interaction of gender with other variables to see why those women were shot and killed.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=race),color="white") + geom_text(stat='count', aes(x=race,label=..count..), vjust=-1) + ggtitle("Plot of Race")
```

Race - Thas was the only variable that truly suprised me when I first looked at it. The news today focuses so much on the killing of black men by police officers after Michael Brown and Trayvon Martin, yet the vast majority of victims according to this data set are white. About 1/3 of the data set are white victims, with black victims coming in second by about 900 less deaths. Granted, this data set does not take into account deaths in custody, shootings from off-duty cops, or shootings that did not result in deaths, but it's interesting that the data tells a slightly different story than the media.

```{r}
ggplot (data=shooting) + geom_bar(mapping=aes(x=city),color="white") + ggtitle("Plot of City of Victim")
```

City - This variable is fairly useless on it's own. There are 1,876 different values for this variable and there is also no consistency. One value that I obsereved is a street address, with others being counties and townships rather than a city itself. I would get rid of this variable and not use it in any bivariate or multivariate explorations.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=state),color="white") + geom_text(stat='count', aes(x=state,label=..count..), vjust=-1) +
  ggtitle("Plot of State of Victim")
```

State - This variable is much more informative than the city variable. What's not suprising to me is that the states with the largest populations seem to have the highest death rates, except for New York. I expected NY to have a much higher police fatality rate due to it's size and population, but then again the majority of the population is concentrated in one city with a police force that is highly trained compared to other police forces in other major cities.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=signs_of_mental_illness),color="white") + geom_text(stat='count', aes(x=signs_of_mental_illness,label=..count..), vjust=-1) +
  ggtitle("Plot of Signs of Mental Illness")
```

Signs of Mental Illness - Another interesting variable to consider, since the news tends to focus on a victim's mental state when it comes to recent police shootings. Like gender, this would be an interesting variable to explore in relation to other variables.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=threat_level),color="white") + geom_text(stat='count', aes(x=threat_level,label=..count..), vjust=-1) + ggtitle("Plot of Threat Level")
```

Threat Level - Personally I would have liked this variable to be classified differently. I feel that other is a very ambiguous term when it comes to threat level. I would convert this varaible to a numeric scale. 5 could be that the victim was armed and actively returning fire, causing a threat to police and civilians in the area. A 0 or 1 could be that the victim is unarmed or fleeing.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=flee),color="white") + geom_text(stat='count', aes(x=flee,label=..count..), vjust=-1) + ggtitle("Plot of Fleeing")
```

Flee - I am surprised that more individuals were not actively fleeing police. I may not be a criminal, but I feel like if I were to commit some sort of crime I would try my best to get away from the police rather than trying to stand my ground. I guess this is why I am not a criminal, as about 2/3 of the victims did not flee, which is surprising to me.

```{r fig.height=10, fig.width=10}
ggplot (data=shooting) + geom_bar(mapping=aes(x=body_camera),color="white") + geom_text(stat='count', aes(x=body_camera,label=..count..), vjust=-1) +
  ggtitle("Plot of Presence of Body Camera")
```

Body Camera - I am actually surprised that the number of incidents where a body camera was present is not higher. I remember that issue being a huge point of conversation surrounding the recent invidents involving police shooting fatalities, so I would have expected the rate to increase a bit more. This would be an intersting relationship to explore along with the date of death to see if there is a rising trend in body camera use.



2) Exploration of Bivariate Relationships

```{r fig.height=15, fig.width=15}
ggplot (data=shooting,mapping=aes(x=gender,fill=threat_level),color="black") + geom_bar(position="dodge") +
  ggtitle("Plot of Threat Level by Gender")
```

Plot of Threat Level by Gender - I mostly chose this graph becasuse I was curious to see the proportion of females that were seen as aggressive/attacking. I was surprised that the proportions seem similar, with about 2/3 of both males and females seen as attacking/aggressive.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting) + geom_point(mapping=aes(x=date,y=body_camera)) + ggtitle("Plot of Body Camera vs Date of Death")
```

Plot of Body Camera vs. Date of Death - As mentioned above, I was curious to see if there was an increase in the presence of body cameras over time, but to me it looks like the presence of body cameras has always been consistently low, even with the concentration of body cameras around early 2016.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting,mapping=aes(x=threat_level,fill=signs_of_mental_illness),color="black") + geom_bar(position="dodge") +
  ggtitle("Plot of Threat Level by Signs of Mental Illness")
```

Plot of Threat Level by Signs of Mental Illness - I chose to graph this relationship to examine if there was any correlation between a vistim's mental state and their perceived threat level. Proportionally, threat level does not seem to be affected that much by the victim's perceived state of mental illness, but there were about twice the amount of individuals that had a perceived state of mental illness as true and were percevied as aggressive/attacking than there were perceived as something as other than attack.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting,mapping=aes(x=body_camera,fill=race),color="black") + geom_bar(position="dodge") +
  ggtitle("Plot of Presence of Body Camera by Race")
```

Plot of Body Camera by Race - Again, I wanted to explore how certain variables tie into race given the media's portrayal and focus on race in today's society. Are there more body camera's used when it comes to a certain race compared to lack of body cameras? Turns out that when a body camera is present there are almost as many black victims as there are white victims, but only about half the total when a body camera is not present.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting,mapping=aes(x=threat_level,fill=race),color="black") + geom_bar(position="dodge") + 
  ggtitle("Plot of Threat Level by Race")
```

Plot of Threat Level by Race - Adding to my exploration of how race factors into police fatalities, I wanted to look at if a certain race was perceived as more aggressive than not, and as it turns out, the answer is really no. More white people are perceived as attacking and aggressive, but again there are a lot more white victims than any other races so proportinally they're about even with the other races in terms of how they're perceived.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting2,mapping=aes(x=armed,fill=threat_level),color="black") + geom_bar(position="dodge") + 
  ggtitle("Plot of Armed by Threat Level")
```

Plot of Armed by Threat Level - To me this graph makes a lot of sense logically. If an officer is faced with someone using a gun, they would typically perceive the other person as attacking and aggressive, and if they weapon they have is only a toy that is hard to determine at first and can easily be confused for a real weapon.


3) Exploration of Multivariate Relationships


```{r fig.height=15, fig.width=15}
ggplot (data=shooting2) + geom_boxplot(mapping=aes(x=armed,y=age,fill=threat_level)) + ggtitle("Plot of Armed, Age, and Threat Level")
```

Plot of Armed, Age, and Threat Level - An interesting thing to note about this plot is that regardless of the weapon used (or no weapon at all), the majority of the victims that were seen as attacking were roughly between the ages of 30 and 40. You can also see that a lot of the unarmed victims trend towards being younger than victims that were carrying some sort of weapon.

```{r fig.height=15, fig.width=15}
ggplot (data=shooting) + geom_point(mapping=aes(x=state,y=age,color=race)) + ggtitle("Plot of State, Age, and Race")
```

Plot of State, Age, and Race - There seems to be a concentration of younger Hispanic victims in the southwestern states like California and Arizona. Southern states see a higher concentration of younger black victims, specifially in Georgia, South Carolina, and Texas

```{r fig.height=15, fig.width=15}
ggplot (data=shooting) + geom_point(mapping=aes(x=date,y=age,color=race)) + ggtitle("Plot of Date, Age, and Race")
```

Plot of Date, Age, and Race - The most interesting thing with this plot is that the older victims over the years tend to mostly be white, whereas the youngest victims over time are predominantly black and hispanic.


4) Exploration of a Multivariate Relationship Using Faceting

```{r fig.height=5}
ggplot (data=shooting) + geom_point(mapping=aes(x=gender,y=age)) + ggtitle("Plot of Gender by Age Faceted by Race") + facet_grid (. ~ race)
```

Gender and Age, Faceted by Race - I chose this graph to facet by race because of the popular media narrative of the "Young Black Man" as being the victim of police brutality. Regardless of where you stand on the issues of racism and police brutality, looking at this data alone you can see that the distribtuion of fatalities among men, by age, is fairly homogenoues across all the races for which there are a larger number of data points. There does not appear to be a strong concentration of young black men being shot and killed by police, or even young men in general. There is a fairly even spread of men of all races and ages that are victims of police shooting. Females on the other hand vary greatly, but mostly because there are significantly less data points, meaning that there are very few women that are fatally shot by police.


5) My Favorite Graph

```{r fig.height=15, fig.width=10}
ggplot (data=shooting,mapping=aes(x=threat_level,fill=race),color="black") + geom_bar(position="dodge") +
  scale_fill_discrete(name="Race", breaks=c("A", "B", "H","N","O","W","NA"),labels=c("Asian", "Black", "Hispanic","Native American","Other","White","N/A")) +
  ggtitle("Plot of Threat Level by Race") +
  labs(x="Threat Level of Victim",y="Count")
```

I chose this plot because I wanted to explore the relationship between the perception of a victim's threat level and their race. Again, revisiting the popular narrative in the media of the aggressive black man committing crimes, I was curious to see if there would be a disproportionate number of blacks that were perceived as "aggressive" or "attacking" police officers. I was surpised to find that, proportionally, there isn't that much of a difference in the data among the races, more specifically between white and black victims as those were the two most prominent races amongst the victims. Just the fact that the number of white deaths almost doubles the number of black deaths goes against this notion of blacks as victim's of police brutality that the media likes to portray. Approximately 68% of victims, despite their race, were viewed with an attacking threat level.

6) Conclusion

Overall, I would say that this data set has a few flaws that prevent it from being a viable source for data analysis. The missing values, the lack of consistency in data categorizing amonst the variables, and the lack of other information and variables prevent me from feeling confident enought to use this data set in a true analysis of police fatalaties in the United States since 2015.